Statistical Guarantees for Local Spectral Clustering on Random Neighborhood Graphs

### 2021-09-10 jmlr/revision2/appendix.tex
Completed edit of paper and appendices. Added a brief response letter.

\section{Population functionals for density clusters}
\label{apdx:density_cluster_population_functionals}
In this appendix, we prove Lemma~\ref{lem:density_cluster_local_spread} (in Section~\ref{subsec:density_cluster_local_spread}), Proposition~\ref{prop:density_cluster_normalized_cut} (in Section~\ref{subsec:density_cluster_ncut}), and Proposition~\ref{prop:density_cluster_conductance} (in Section~\ref{subsec:density_cluster_conductance}), by establishing bounds on the population-level local spread, normalized cut, and conductance of a thickened density cluster $\mc{C}_{\lambda,\sigma}$. In these proofs, we make use of some estimates on the volume of spherical caps (in Section~\ref{subsec:spherical_caps}); some isoperimetric inequalities (in Section~\ref{subsec:isoperimetric_inequalities}), and some reverse isoperimetric inequalities (given in Section~\ref{subsec:reverse_isoperimetric_inequalities}). Finally, in Section~\ref{subsec:density_cluster_hard_case}, for the hard case distribution $\Pbb$ defined in~\eqref{eqn:lb_density} and $\mc{L}$ defined in~\eqref{eqn:lower_set}, establish bounds on the population-level normalized cut $\Phi_{\Pbb,r}(\mc{L})$ and local spread $s_{\Pbb,r}(\mc{X})$; these will be useful in the proof of Theorem~\ref{thm:ppr_lb}. Throughout, we write $\nu_d := \nu(B(0,1))$ for the Lebesgue measure of a $d$-dimensional unit ball.

\subsection{Balls, Spherical Caps, and Associated Estimates}
\label{subsec:spherical_caps}
In this section, we derive lower bounds on the volume of the intersection between two balls in $\Rd$, and the volume of a spherical cap. Results of this type are well-known, but since we could not find exactly the statements we desire, for completeness we also supply proofs. We use the notation $B(x,r)$ for a ball of radius $r$ centered at $x \in \Rd$, and $\mathrm{cap}_{r}(h)$ for a spherical cap of height $r$ and radius $r$. Recall that the Lebesgue measure of a spherical cap is
\begin{equation*}
\nu\bigl(\mathrm{cap}_r(h)\bigr) = \frac{1}{2} \nu_d r^d I_{1 - a}\left(\frac{d + 1}{2}; \frac{1}{2}\right)
\end{equation*}
where $a = (r - h)^2/r^2$, and
\begin{equation*}
I_{1 - a}(z,w) = \frac{\Gamma(z + w)}{\Gamma(z) \Gamma(w)} \int_{0}^{1 - a} u^{z - 1} (1 - u)^{w - 1} du.
\end{equation*}
is the cumulative distribution function of a $\mathrm{Beta}(z,w)$ distribution, evaluated at $1 - a$. (Here $\Gamma(\cdot)$ is the gamma function).
\begin{lemma}
	\label{lem:overlap_balls}
	For any $x,y \in \Rd$ and $r > 0$, it holds that
	\begin{equation}
	\label{eqn:overlap_balls_1}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d\biggl(1 - \frac{\|x - y\|}{r} \sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
	For any $x,y \in \Rd$ and $r,\sigma > 0$ such that $\|x - y\| \leq \sigma$, it holds that,
	\begin{equation}
	\label{eqn:overlap_balls_2}
	\nu\bigl(B(x,r) \cap B(y,\sigma)\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
\end{lemma}
\begin{lemma}
	\label{lem:volume_of_spherical_cap}
	For any $0 < h \leq r$, and $a = 1 - (2 r h - h^2)/r^2$,
	\begin{equation*}
	\nu\bigl(\mathrm{cap}_r(h)\bigr) \geq \frac{1}{2}\nu_dr^d\bigl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\bigr)
	\end{equation*}
\end{lemma}
An immediate implication of~\eqref{eqn:overlap_balls_2} is that for any $x \in \mc{C}_{\lambda,\sigma}$,
\begin{equation}
\label{eqn:uniform_local_conductance}
\nu\bigl(B(x,r) \cap \mc{C}_{\lambda,\sigma}\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
\end{equation}  
\paragraph{Proof of Lemma~\ref{lem:overlap_balls}.}
	First, we prove~\eqref{eqn:overlap_balls_1}. The intersection $B(x,r) \cap B(y,r)$ consists of two symmetric spherical caps, each with height $h = r - \frac{\|x - y\|}{2}$. 
	As a result, by Lemma~\ref{lem:volume_of_spherical_cap} we have
	\begin{equation*}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d \bigl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\bigr)
	\end{equation*}
	where $a = \|x - y\|^2/(4r^2)$, and the claim follows.

@@

\subsection{Isoperimetric inequalities}
\label{subsec:isoperimetric_inequalities}
\citet{dyer1991} establish the following isoperimetric inequality for convex sets.
\begin{lemma}[Isoperimetry of a convex set.]
	\label{lem:convex_isoperimetric_inequality}
	For any partition $(\mc{R}_1,\mc{R}_2,\mc{R}_3)$ of a convex set $\mc{K} \subseteq \Rd$, it holds that
	\begin{equation*}
	\nu(\mc{R}_3) \geq 2\frac{\mathrm{dist}(\mc{R}_1, \mc{R}_2)}{\mathrm{diam}(\mc{K})} \min(\nu(\mc{R}_1), \nu(\mc{R}_2)).
	\end{equation*}
\end{lemma}
\citet{abbasi-yadkori2016a} points out that if $\mc{S}$ is the image of a convex set under a Lipschitz measure-preserving mapping $g: \Rd \to \Rd$, a similar inequality can be obtained.
\begin{corollary}[Isoperimetry of Lipschitz embeddings of convex sets.]
	\label{cor:nonconvex_isoperimetric_inequality}
	Suppose $\mc{S}$ is the image of a convex set $\mathcal{K}$ under a mapping $g:\Rd \to \Rd$ such that
	\begin{equation*}
	\|g(x) - g(y)\| \leq M \cdot \|x - y\|,~~\textrm{for all $x,y \in \mc{K}$, and}~~\det(\nabla g(x)) = 1~~\textrm{for all $x \in \mc{K}$.}
	\end{equation*}
	Then for any partition $(\Omega_1,\Omega_2,\Omega_3)$ of $\mc{S}$, 
	\begin{equation*}
	\nu(\Omega_3) \geq 2\frac{\mathrm{dist}(\Omega_1, \Omega_2)}{\diam(\mc{K}) M} \min(\nu(\Omega_1), \nu(\Omega_2)).
	\end{equation*}
\end{corollary}

@@

	It remains to verify that $s \mapsto \nu\bigl(\mc{A}_j(s)\bigr)/\nu\bigl(B(x_j,s)\bigr)$ is monotonically non-increasing. For any $0 < s < t$ and $j = 1,\ldots,N$, suppose $x \in \mc{A}_j(T) - \{x_j\}$, meaning $x \in B(0,t)$ and $x \not\in B(x_i - x_j,t)$ for any $i = 1,\ldots,j - 1$. Thus $(s/t)x \in B(0,s)$, and
	\begin{equation*}
	\|(s/t)x - (x_i - x_j)\| \geq \|x - (x_i - x_j)\| - \|x - (s/t)x\| > t - (1 - s/t)\|x\| \geq t - (1 - s/t)t = s,
	\end{equation*}
	or in other words $(s/t)x \not\in B(x_i - x_j,s)$ for any $i = 1,\ldots,j - 1$. Consequently,
	\begin{equation*}
	\Bigl(\mc{A}_j(t) - \{x_j\}\Bigr) \subset \frac{t}{s} \cdot \Bigl(\mc{A}_{j}(s)- \{x_j\}\Bigr),
	\end{equation*}
	and applying $\nu(\cdot)$ to both sides yields the claim.

\paragraph{Proof of Lemma~\ref{lem:reverse_isoperimetric_inequality_density_weighted}.}
	Fix $k \in \mathbb{N}$. To establish~\eqref{eqn:reverse_isoperimetric_inequality_density_weighted}, we partition $\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}$ into thin tubes $\mc{T}_{1},\ldots,\mc{T}_{k}$, with the $j$th tube $\mc{T}_j$ defined as $\mc{T}_j := \mc{C}_{\lambda,\sigma + jr/k} \setminus \mc{C}_{\lambda,\sigma + (j - 1)r/k}.$ We upper bound the Lebesgue measure of each tube $\mc{T}_j$ using~\eqref{eqn:reverse_isoperimetric_inequality_2}:\footnote{Note that $\mc{C}$ must be bounded, since the density $f(x) \geq \lambda_{\sigma}$ for all $x \in \mc{C}$.}
	\begin{equation*}
	\nu(\mc{T}_j) \leq \frac{dr/k}{\sigma - dr/k} \nu(\mc{C}_{\lambda,\sigma + (j - 1)r/k}) \leq \frac{dr/k}{\sigma - dr/k} \nu(\mc{C}_{\lambda,\sigma + r}) \leq \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot \frac{dr/k}{\sigma - dr/k}  \cdot \nu(\mc{C}_{\lambda,\sigma}),
	\end{equation*}
	and the maximum density within each tube using~\ref{asmp:low_noise_density}:
	\begin{equation*}
	\max_{x \in \mc{T}_j} f(x) \leq \lambda_{\sigma} - \theta \Bigl(\frac{j - 1}{k}r\Bigr)^{\gamma};
	\end{equation*}
	combining these upper bounds, we see that
	\begin{equation}
	\label{pf:reverse_isoperimetric_inequality_density_weighted_1}
	\Pbb\bigl(\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}\bigr) = \sum_{j = 1}^{k} \Pbb(\mc{T}_j) \leq \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot  \frac{dr/k}{\sigma - dr/k} \cdot \nu(\mc{C}_{\lambda,\sigma}) \cdot \biggl(\sum_{j = 0}^{k - 1}\lambda_{\sigma} -  \theta r^{\gamma} \Bigl(\frac{j}{k}\Bigr)^{\gamma} \biggr).
	\end{equation}
	Treating the sum in the previous expression as a Riemann sum of a non-increasing function evaluated at $0,\ldots,k -1$ gives the upper bound
	\begin{equation*}
	\sum_{j = 0}^{k - 1}\lambda_{\sigma} -  \theta r^{\gamma} \Bigl(\frac{j}{k}\Bigr)^{\gamma } \leq \lambda_{\sigma} + \int_{0}^{k - 1} \Bigl(\lambda_{\sigma} -  \theta r^{\gamma} \Bigl(\frac{x}{k}\Bigr)^{\gamma}\Bigr) \,dx \leq k\lambda_{\sigma} + (k - 1)\frac{\theta r^{\gamma}}{\gamma + 1} \Bigl(\frac{k - 1}{k}\Bigr)^{\gamma},
	\end{equation*}
	and plugging back in to~\eqref{pf:reverse_isoperimetric_inequality_density_weighted_1}, we obtain
	\begin{equation*}
	\Pbb\bigl(\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}\bigr) \leq  \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot \frac{dr}{\sigma - dr/k} \nu(\mc{C}_{\lambda,\sigma}) \cdot \biggl(\lambda - \frac{\theta r^{\gamma}}{\gamma + 1} \cdot \Bigl(\frac{k - 1}{k}\Bigr)^{\gamma + 1} \biggl).
	\end{equation*}
	The above inequality holds for any $k \in \mathbb{N}$, and taking the limit of the right hand side as $k \to \infty$ yields the claim.

\subsection{Proof of Lemma~\ref{lem:density_cluster_local_spread}}
\label{subsec:density_cluster_local_spread}
The population-level local spread of $\mc{C}_{\lambda,\sigma}$ is
\begin{equation*}
s_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) = \frac{\bigl(d_{\min}(\wt{\Pbb})\bigr)^2}{\vol_{\wt{\Pbb},r}(\mc{C}_{\lambda},\sigma)}
\end{equation*}
where we recall that $\wt{\Pbb}(\mc{S}) = \frac{\Pbb(\mc{S} \cap \mc{C}_{\lambda,\sigma})}{\Pbb(\mc{C}_{\lambda,\sigma})}$ for Borel sets $\mc{S}$, and $d_{\min}(\wt{\Pbb}) := \min_{x \in \mc{C}_{\lambda,\sigma}}\{\deg_{\wt{\Pbb},r}(x)\}^2$. To lower bound $s_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$, we first lower bound $d_{\min}(\wt{\Pbb})$, and then upper bound $\vol_{\wt{\Pbb},r}(\mc{C}_{\lambda},\sigma)$. Using the lower bound $f(x) \geq \lambda_{\sigma}$ for all $x \in \mc{C}_{\lambda,\sigma}$ stipulated in~\ref{asmp:lambda_bounded_density}, we deduce that
\begin{align*}
d_{\min}(\wt{\Pbb}) & = \min_{x \in \mc{C}_{\lambda,\sigma}} \Bigl\{ \int \1\{\|x - y\| \leq r\} \,d\wt{\Pbb}(y) \Bigr\} \\
& \geq \frac{\lambda_{\sigma}}{\Pbb(\mc{C}_{\lambda,\sigma})} \cdot \min_{x \in \mc{C}_{\lambda,\sigma}} \Bigl\{ \int_{\mc{C}_{\lambda,\sigma}} \1\{\|x - y\| \leq r\} \,dy \Bigr\} \\
& \geq \frac{\lambda_{\sigma}}{\Pbb(\mc{C}_{\lambda,\sigma})} \cdot \frac{1}{2}\nu_dr^d \cdot \Bigl(1 - \frac{r}{\sigma} \sqrt{\frac{d + 2}{2\pi}}\Bigr),
\end{align*}
where the final inequality follows from Lemma~\ref{lem:overlap_balls}.

On the other hand, using the upper bound $f(x) \leq \Lambda_{\sigma}$ for all $x \in \mc{C}_{\lambda,\sigma}$, we deduce that
\begin{align*}
\vol_{\wt{\Pbb},r}(\mc{C}_{\lambda,\sigma}) & = \iint \1\{\|x - y\|\leq r\} \,d\wt{\Pbb}(y) \,d\wt{\Pbb}(x) \\
& \leq \frac{\Lambda_{\sigma}^2}{\Pbb(\mc{C}_{\lambda,\sigma})^2} \cdot \int_{\mc{C}_{\lambda,\sigma}} \int_{\mc{C}_{\lambda,\sigma}} \1\{\|x - y\|\leq r\} \,dy \,dx \\
& \leq \frac{\Lambda_{\sigma}^2}{\Pbb(\mc{C}_{\lambda,\sigma})^2} \cdot \nu_d r^d \cdot \nu(\mc{C}_{\lambda,\sigma}) \\
& \leq \frac{\Lambda_{\sigma}^2}{\Pbb(\mc{C}_{\lambda,\sigma})^2} \cdot \nu_d^2 r^d  \cdot \Bigl(\frac{\rho}{2}\Bigr)^d;
\end{align*}
the final inequality follows from~\ref{asmp:embedding}, which implies that $\nu(\mc{C}_{\lambda,\sigma}) = \nu(\mc{K}) \leq \nu_d(\rho/2)^d$. The claim of Lemma~\ref{lem:density_cluster_local_spread} follows.

@@

	As is standard (see for example~\cite{dyer1991,lovasz1990}), the proof of Lemma~\ref{lem:uniform_density_cluster_conductance} uses several inequalities to lower bound the normalized cut $\Phi_{\wt{\nu},r}(\mc{S})$.

@@

	We turn to lower bounding $\cut_{n,r}(Z)$, which is considerably more involved. We will approximate the cut of $Z$ by discretizing the space $\mathcal{X}$ into bins, relate the cut of $Z$ to the boundary of the binned set $\overline{Z}$, and then lower bound the size of the boundary of $\overline{Z}$.

>>>>

\section{Population functionals for density clusters}
\label{apdx:density_cluster_population_functionals}
In this appendix, we prove Lemma~\ref{lem:density_cluster_local_spread} (in Section~\ref{subsec:density_cluster_local_spread}), Proposition~\ref{prop:density_cluster_normalized_cut} (in Section~\ref{subsec:density_cluster_ncut}), and Proposition~\ref{prop:density_cluster_conductance} (in Section~\ref{subsec:density_cluster_conductance}), by establishing bounds on the population-level local spread, normalized cut, and conductance of a thickened density cluster $\mc{C}_{\lambda,\sigma}$. In these proofs, we make use of some estimates on the volume of spherical caps (given in Section~\ref{subsec:spherical_caps}); some isoperimetric inequalities (Section~\ref{subsec:isoperimetric_inequalities}), and some reverse isoperimetric inequalities (Section~\ref{subsec:reverse_isoperimetric_inequalities}). Finally, in Section~\ref{subsec:density_cluster_hard_case}, for the hard case distribution $\Pbb$ defined in~\eqref{eqn:lb_density} and $\mc{L}$ defined in~\eqref{eqn:lower_set}, we establish bounds on the population-level normalized cut $\Phi_{\Pbb,r}(\mc{L})$ and local spread $s_{\Pbb,r}(\mc{X})$; these will be useful in the proof of Theorem~\ref{thm:ppr_lb}. Throughout, we write $\nu_d := \nu(B(0,1))$ for the Lebesgue measure of a $d$-dimensional unit ball.

\subsection{Balls, Spherical Caps, and Associated Estimates}
\label{subsec:spherical_caps}
In this section, we derive lower bounds on the volume of the intersection between two balls in $\Rd$, and the volume of a spherical cap. Results of this type are well-known, but since we could not find exactly the statements we desire, for completeness we also supply proofs. We use the notation $B(x,r)$ for a ball of radius $r$ centered at $x \in \Rd$, and $\mathrm{cap}_{r}(h)$ for a spherical cap of height $h$ and radius $r$. Recall that the Lebesgue measure of a spherical cap is
\begin{equation*}
\nu\bigl(\mathrm{cap}_r(h)\bigr) = \frac{1}{2} \nu_d r^d I_{1 - a}\left(\frac{d + 1}{2}; \frac{1}{2}\right),
\end{equation*}
where $a = (r - h)^2/r^2$, and
\begin{equation*}
I_{1 - a}(z,w) = \frac{\Gamma(z + w)}{\Gamma(z) \Gamma(w)} \int_{0}^{1 - a} u^{z - 1} (1 - u)^{w - 1} du,
\end{equation*}
is the cumulative distribution function of a $\mathrm{Beta}(z,w)$ distribution, evaluated at $1 - a$. (Here $\Gamma(\cdot)$ is the gamma function).
\begin{lemma}
	\label{lem:overlap_balls}
	For any $x,y \in \Rd$ and $r > 0$, it holds that
	\begin{equation}
	\label{eqn:overlap_balls_1}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d\biggl(1 - \frac{\|x - y\|}{r} \sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
	For any $x,y \in \Rd$ and $r,\sigma > 0$ such that $\|x - y\| \leq \sigma$, it holds that,
	\begin{equation}
	\label{eqn:overlap_balls_2}
	\nu\bigl(B(x,r) \cap B(y,\sigma)\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation}
\end{lemma}
\begin{lemma}
	\label{lem:volume_of_spherical_cap}
	For any $0 < h \leq r$, and $a = 1 - (2 r h - h^2)/r^2$,
	\begin{equation*}
	\nu\bigl(\mathrm{cap}_r(h)\bigr) \geq \frac{1}{2}\nu_dr^d\biggl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\biggr).
	\end{equation*}
\end{lemma}
An immediate implication of~\eqref{eqn:overlap_balls_2} is that for any $x \in \mc{C}_{\lambda,\sigma}$,
\begin{equation}
\label{eqn:uniform_local_conductance}
\nu\bigl(B(x,r) \cap \mc{C}_{\lambda,\sigma}\bigr) \geq \frac{1}{2} \nu_d r^d\biggl(1 - \frac{r}{\sigma}\sqrt{\frac{d + 2}{2\pi}}\biggr).
\end{equation}  
\paragraph{Proof of Lemma~\ref{lem:overlap_balls}.}
	First, we prove~\eqref{eqn:overlap_balls_1}. The intersection $B(x,r) \cap B(y,r)$ consists of two symmetric spherical caps, each of height $h = r - \frac{\|x - y\|}{2}$. 
	As a result, by Lemma~\ref{lem:volume_of_spherical_cap} we have
	\begin{equation*}
	\nu\bigl(B(x,r) \cap B(y,r)\bigr) \geq \nu_d r^d \bigl(1 - 2\sqrt{a} \cdot \sqrt{\frac{d + 2}{2\pi}}\bigr)
	\end{equation*}
	where $a = \|x - y\|^2/(4r^2)$, and the claim follows.

@@

\subsection{Isoperimetric inequalities}
\label{subsec:isoperimetric_inequalities}
\citet{dyer1991b} establish the following isoperimetric inequality for convex sets.
\begin{lemma}[Isoperimetry of a convex set.]
	\label{lem:convex_isoperimetric_inequality}
	For any partition $(\mc{R}_1,\mc{R}_2,\mc{R}_3)$ of a convex set $\mc{K} \subseteq \Rd$, it holds that
	\begin{equation*}
	\nu(\mc{R}_3) \geq 2\frac{\mathrm{dist}(\mc{R}_1, \mc{R}_2)}{\mathrm{diam}(\mc{K})} \min(\nu(\mc{R}_1), \nu(\mc{R}_2)).
	\end{equation*}
\end{lemma}
\citet{abbasi-yadkori2016a} points out that if $\mc{S}$ is the image of a convex set under a Lipschitz measure-preserving mapping $g: \Rd \to \Rd$, a similar inequality can be obtained.
\begin{corollary}[Isoperimetry of Lipschitz embeddings of convex sets.]
	\label{cor:nonconvex_isoperimetric_inequality}
	Suppose $\mc{S}$ is the image of a convex set $\mathcal{K}$ under a mapping $g:\Rd \to \Rd$ such that
	\begin{equation*}
	\|g(x) - g(y)\| \leq M \cdot \|x - y\|,~~\textrm{for all $x,y \in \mc{K}$, and}~~\det(\nabla g(x)) = 1~~\textrm{for all $x \in \mc{K}$.}
	\end{equation*}
	Then for any partition $(\Omega_1,\Omega_2,\Omega_3)$ of $\mc{S}$, 
	\begin{equation*}
	\nu(\Omega_3) \geq 2\frac{\mathrm{dist}(\Omega_1, \Omega_2)}{\diam(\mc{K}) M} \min(\nu(\Omega_1), \nu(\Omega_2)).
	\end{equation*}
\end{corollary}

@@

	It remains to verify that $s \mapsto \nu\bigl(\mc{A}_j(s)\bigr)/\nu\bigl(B(x_j,s)\bigr)$ is monotonically non-increasing. For any $0 < s < t$ and $j = 1,\ldots,N$, suppose $x \in \mc{A}_j(T) - x_j$, meaning $x \in B(0,t)$ and $x \not\in B(x_i - x_j,t)$ for any $i = 1,\ldots,j - 1$. Thus $(s/t)x \in B(0,s)$, and
	\begin{equation*}
	\|(s/t)x - (x_i - x_j)\| \geq \|x - (x_i - x_j)\| - \|x - (s/t)x\| > t - (1 - s/t)\|x\| \geq t - (1 - s/t)t = s,
	\end{equation*}
	or in other words $(s/t)x \not\in B(x_i - x_j,s)$ for any $i = 1,\ldots,j - 1$. Consequently,
	\begin{equation*}
	\Bigl(\mc{A}_j(t) - x_j\Bigr) \subset \frac{t}{s} \cdot \Bigl(\mc{A}_{j}(s)- x_j\Bigr),
	\end{equation*}
	and applying $\nu(\cdot)$ to both sides yields the claim.

\paragraph{Proof of Lemma~\ref{lem:reverse_isoperimetric_inequality_density_weighted}.}
	Fix $k \in \mathbb{N}$. To establish~\eqref{eqn:reverse_isoperimetric_inequality_density_weighted}, we partition $\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}$ into thin tubes $\mc{T}_{1},\ldots,\mc{T}_{k}$, with the $j$th tube $\mc{T}_j$ defined as $\mc{T}_j := \mc{C}_{\lambda,\sigma + jr/k} \setminus \mc{C}_{\lambda,\sigma + (j - 1)r/k}.$ We upper bound the Lebesgue measure of each tube $\mc{T}_j$ using~\eqref{eqn:reverse_isoperimetric_inequality_2}:\footnote{Note that $\mc{C}$ must be bounded, since the density $f(x) \geq \lambda_{\sigma} > 0$ for all $x \in \mc{C}$.}
	\begin{equation*}
	\nu(\mc{T}_j) \leq \frac{dr/k}{\sigma - dr/k} \nu(\mc{C}_{\lambda,\sigma + (j - 1)r/k}) \leq \frac{dr/k}{\sigma - dr/k} \nu(\mc{C}_{\lambda,\sigma + r}) \leq \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot \frac{dr/k}{\sigma - dr/k}  \cdot \nu(\mc{C}_{\lambda,\sigma}),
	\end{equation*}
	and the maximum density within each tube using~\ref{asmp:low_noise_density}:
	\begin{equation*}
	\max_{x \in \mc{T}_j} f(x) \leq \lambda_{\sigma} - \theta \Bigl(\frac{j - 1}{k}r\Bigr)^{\gamma};
	\end{equation*}
	combining these upper bounds, we see that
	\begin{equation}
	\label{pf:reverse_isoperimetric_inequality_density_weighted_1}
	\Pbb\bigl(\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}\bigr) = \sum_{j = 1}^{k} \Pbb(\mc{T}_j) \leq \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot  \frac{dr/k}{\sigma - dr/k} \cdot \nu(\mc{C}_{\lambda,\sigma}) \cdot \biggl(\sum_{j = 0}^{k - 1}\lambda_{\sigma} -  \theta r^{\gamma} \Bigl(\frac{j}{k}\Bigr)^{\gamma} \biggr).
	\end{equation}
	Treating the sum in the previous expression as a Riemann sum of a non-increasing function evaluated at $0,\ldots,k -1$ gives the upper bound
	\begin{equation*}
	\sum_{j = 0}^{k - 1}\lambda_{\sigma} -  \theta r^{\gamma} \Bigl(\frac{j}{k}\Bigr)^{\gamma } \leq \lambda_{\sigma} + \int_{0}^{k - 1} \Bigl(\lambda_{\sigma} -  \theta r^{\gamma} \Bigl(\frac{x}{k}\Bigr)^{\gamma}\Bigr) \,dx \leq k\lambda_{\sigma} + (k - 1)\frac{\theta r^{\gamma}}{\gamma + 1} \Bigl(\frac{k - 1}{k}\Bigr)^{\gamma},
	\end{equation*}
	and plugging back in to~\eqref{pf:reverse_isoperimetric_inequality_density_weighted_1}, we obtain
	\begin{equation*}
	\Pbb\bigl(\mc{C}_{\lambda,\sigma + r} \setminus \mc{C}_{\lambda,\sigma}\bigr) \leq  \Bigl(1 + \frac{dr}{\sigma - dr}\Bigr) \cdot \frac{dr}{\sigma - dr/k} \nu(\mc{C}_{\lambda,\sigma}) \cdot \biggl(\lambda - \frac{\theta r^{\gamma}}{\gamma + 1} \cdot \Bigl(\frac{k - 1}{k}\Bigr)^{\gamma + 1} \biggl).
	\end{equation*}
	The above inequality holds for any $k \in \mathbb{N}$, and taking the limit of the right hand side as $k \to \infty$ yields the claim.

\subsection{Proof of Lemma~\ref{lem:density_cluster_local_spread}}
\label{subsec:density_cluster_local_spread}
The population-level local spread of $\mc{C}_{\lambda,\sigma}$ is
\begin{equation*}
s_{\Pbb,r}(\mc{C}_{\lambda,\sigma}) = \frac{\bigl(d_{\min}(\wt{\Pbb})\bigr)^2}{\vol_{\wt{\Pbb},r}(\mc{C}_{\lambda},\sigma)},
\end{equation*}
where we recall that $\wt{\Pbb}(\mc{S}) = \frac{\Pbb(\mc{S} \cap \mc{C}_{\lambda,\sigma})}{\Pbb(\mc{C}_{\lambda,\sigma})}$ for Borel sets $\mc{S}$, and $d_{\min}(\wt{\Pbb}) := \min_{x \in \mc{C}_{\lambda,\sigma}}\{\deg_{\wt{\Pbb},r}(x)\}^2$. To lower bound $s_{\Pbb,r}(\mc{C}_{\lambda,\sigma})$, we first lower bound $d_{\min}(\wt{\Pbb})$, and then upper bound $\vol_{\wt{\Pbb},r}(\mc{C}_{\lambda},\sigma)$. Using the lower bound $f(x) \geq \lambda_{\sigma}$ for all $x \in \mc{C}_{\lambda,\sigma}$ stipulated in~\ref{asmp:lambda_bounded_density}, we deduce that
\begin{align*}
d_{\min}(\wt{\Pbb}) & = \min_{x \in \mc{C}_{\lambda,\sigma}} \Bigl\{ \int \1\{\|x - y\| \leq r\} \,d\wt{\Pbb}(y) \Bigr\} \\
& \geq \frac{\lambda_{\sigma}}{\Pbb(\mc{C}_{\lambda,\sigma})} \cdot \min_{x \in \mc{C}_{\lambda,\sigma}} \Bigl\{ \int_{\mc{C}_{\lambda,\sigma}} \1\{\|x - y\| \leq r\} \,dy \Bigr\} \\
& \geq \frac{\lambda_{\sigma}}{\Pbb(\mc{C}_{\lambda,\sigma})} \cdot \frac{1}{2}\nu_dr^d \cdot \Bigl(1 - \frac{r}{\sigma} \sqrt{\frac{d + 2}{2\pi}}\Bigr),
\end{align*}
where the final inequality follows from Lemma~\ref{lem:overlap_balls}.

On the other hand, using the upper bound $f(x) \leq \Lambda_{\sigma}$ for all $x \in \mc{C}_{\lambda,\sigma}$, we deduce that
\begin{align*}
\vol_{\wt{\Pbb},r}(\mc{C}_{\lambda,\sigma}) & = \iint \1\{\|x - y\|\leq r\} \,d\wt{\Pbb}(y) \,d\wt{\Pbb}(x) \\
& \leq \frac{\Lambda_{\sigma}^2}{\Pbb(\mc{C}_{\lambda,\sigma})^2} \cdot \int_{\mc{C}_{\lambda,\sigma}} \int_{\mc{C}_{\lambda,\sigma}} \1\{\|x - y\|\leq r\} \,dy \,dx \\
& \leq \frac{\Lambda_{\sigma}^2}{\Pbb(\mc{C}_{\lambda,\sigma})^2} \cdot \nu_d r^d \cdot \nu(\mc{C}_{\lambda,\sigma}) \\
& \leq \frac{\Lambda_{\sigma}^2}{\Pbb(\mc{C}_{\lambda,\sigma})^2} \cdot \nu_d^2 r^d  \cdot \Bigl(\frac{\rho}{2}\Bigr)^d;
\end{align*}
the final inequality follows from~\ref{asmp:embedding}, which along with the isodiametric inequality for convex sets~\citep{gruber2007} implies $\nu(\mc{C}_{\lambda,\sigma}) = \nu(\mc{K}) \leq \nu_d(\rho/2)^d$. The claim of Lemma~\ref{lem:density_cluster_local_spread} follows.

@@

	As is standard (see for example~\cite{dyer1991b,lovasz1990}), the proof of Lemma~\ref{lem:uniform_density_cluster_conductance} uses several inequalities to lower bound the normalized cut $\Phi_{\wt{\nu},r}(\mc{S})$.

@@

	We turn to lower bounding $\cut_{n,r}(Z)$. Establishing this lower bound is considerably more involved, and we start by giving an outline of our approach to build intuition.
	\begin{itemize}
		\item \emph{Discretization}. Intuitively, we will approximate the cut of $Z$ by discretizing the space $\mathcal{X}$ into a collection of rectangular bins $\wb{Q}$. We will consider a subset of bins $\partial \wb{Z} \subseteq \wb{Q}$, which we call the discretized boundary of $Z$, and which consists of those bins $Q \in \wb{Q}$ containing many points $x_i \in Z$ adjacent, in the graph $G_{n,r}$, to many $x_j \in X \setminus Z$. This discretization allows us to relate $\cut_{n,r}(Z)$ to the number of bins in the discretized boundary $\partial \wb{Z}$, as in~\eqref{eqn:normalized_cut_lb_pf1}.
		\item \emph{Slicing}. In order to lower bound the number of bins in the discretized boundary $|\partial \wb{Z}|$, we consider a second partition of $\mc{X}$ into horizontal slices $R$. We argue that for each slice $R$, one of two things must be true: either $R$ contains a bin belonging to the discretized boundary of $Z$, $Q \in \partial \wb{Z}$, or $(Z \vartriangle \mc{C}^{(1)}[X]) \cap R$ has a ``substantial'' volume,
		\begin{equation*}
		\vol_{n,r}\Bigl((Z \vartriangle \mc{C}^{(1)}[X]) \cap R\Bigr) \geq R_{\min},
		\end{equation*}
		where $R_{\min}$ is a random quantity defined in~\eqref{eqn:normalized_cut_lb_pf2}.
		Summing over all slices $R$ then gives a lower bound on $|\partial \wb{Z}|$, as in~\eqref{eqn:normalized_cut_lb_pf2}. 
		\item \emph{Probabilistic Guarantees}. The lower bound~\eqref{eqn:normalized_cut_lb_pf2} depends on two random quantities: the minimum number of points in any bin (denoted by $Q_{\min}$), and $R_{\min}$, which was discussed in the previous bullet. We give suitable lower bounds on each quantity, which hold with high probability, and which suffice to complete the proof of Lemma~\ref{lem:normalized_cut_lb}.
	\end{itemize}

	Now, we fill in the details, working step-by-step.
